---
title: "Macular Degeneration Analysis"
format: pdf
editor: visual
---

```{r setup}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
library(tidyverse)
library(vroom)
library(nlme)
library(multcomp)
library(car)

armd <- vroom('ARMD.txt')
```

## EDA

### Violin Plot

It looks like the treatment is effective, but mostly just over longer periods of time.

```{r}
ggplot(data = armd, aes(x=Vision, y=as.factor(Time), fill=as.factor(Trt))) +
  geom_violin()
```

### Line Plot

Vision appears to be worsening much faster without treatment than it is with treatment.

```{r}
armd %>%
  mutate(Time = as.factor(Time),
         Trt = as.factor(Trt)) %>%
  group_by(Trt, Time) %>%
  summarize(Vision = mean(Vision)) %>%
  ggplot(data=., aes(x=Time, y=Vision, color=Trt, group=Trt)) +
  geom_line()
```

### Correlation within Subject

```{r}
ind.lm <- lm(Vision~. -Subject, data=armd)
ind_resids <- matrix(data=ind.lm$residuals, ncol=4, byrow = TRUE)
cor(ind_resids)
```

## Box Regression Model

```{r}
armd.gls <- gls(Vision ~ Baseline + Time*Trt, data = armd,
                 correlation = corSymm(form = ~4|Subject))
coef(armd.gls) # beta-hat
sigma(armd.gls) # sigma-hat
armd.gls$modelStruct$corStruct #correlation parameters
```

### Validate Model

```{r}
armd.lm <- lm(Vision ~., data = armd)
avPlots(armd.lm, terms = ~Baseline+Time, ask = FALSE)

armd$residuals <- residuals(armd.gls, type = 'normalized')
wide_armd <- armd %>%
  dplyr::select(Subject, Time, residuals) %>%
  pivot_wider(names_from = Time, values_from = residuals)
residual_matrix <- as.matrix(wide_armd[,-1])
cor_matrix <- cor(residual_matrix)
cor_matrix

armd %>% 
  ggplot() +
  geom_histogram(aes(x=residuals, y = after_stat(density))) +
  stat_function(fun = dnorm, args = list(mean = 0, sd = 1), color = "red")

armd$fitted_values <- fitted(armd.gls)
armd %>% 
  ggplot(aes(x = fitted_values, y = residuals)) +
  geom_point()
```

## Compare Trt to No Trt

```{r}
summary(armd.gls)
summary(glht(armd.gls, linfct = matrix(c(0,0,0,1,1),nrow=1), alternative='less'))
```

## Confidence Intervals

The interaction confidence interval tells us that as time increases, vision of those with the treatment improves.

Vision generally decreases with time.

Treatment alone could lead to increase or decrease in vision scores. Doesn't tell us much by itself.

```{r}
confint(armd.gls)
```

## Question 4

average vision at 52 weeks

```{r}
armd_52weeks <- armd %>%
  filter(Time == 52)
armd_52weeks %>%
  group_by(Trt) %>%
  summarize(mean_vision = mean(Vision, na.rm = TRUE))
```

There is a significant difference at 52 weeks

```{r}
t.test(Vision ~ Trt, data = armd_52weeks)
```
